### 2019-05-23 reviews/response.md

We thank the reviewers for their detailed and insightful feedback.

We have added three new implementations to Task Bench: Dask,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section.

One of the salient features of Task Bench is that it permits the
implementation of a benchmark with M configurations and N systems with
an amount of effort that is O(M+N) rather than O(MN). We have updated
the introduction to reflect this.

Major changes in the revised paper:

  * Three new implementations (Dask, MPI+OpenMP, MPI+CUDA)
  * New sections
      - Implementation: sections 4.3 (Dask) and 4.5 (MPI+X)
      - Evaluation: sections 5.6 (weak/strong scaling) and 5.9 (GPUs)
  * New figures
      - Listings 1 and 2 (code samples)
      - Figures 11, 12, 15 (experiments)

Reviewer 1:

The memory bound kernel has been written to perform a varying number
of passes (or partial passes) over a fixed-sized buffer, in order to
keep working set size constant and avoid interference due to cache
effects, which in our opinion makes the results difficult to
interpret.

See the code samples in Listing 1 and 2.

Reviewer 2:

Figures 3 and 4 have been updated and new text added.

Reviewer 3:

Weak/strong scaling and TPS are all application and system dependent
metrics. METG differs from these metrics in that can simultaneously
(a) capture the behavior of the sytem at the limit of its scalability
(which is not true of weak/strong scaling) and (b) while achieving a
specified amount of useful work (which is not true of TPS). METG also
isolates the impact of changes to the communication topology at scale,
whereas in a strong scaling study this effect would be entangled with
issues of system overhead. Examples of how to compute METG from
application wall time have been added to Section 2.

METG captures the average task granularity that can be supported. It
is acceptable for some tasks to be smaller and others larger, as
overheads will even out across tasks.

A comparison with weak/strong scaling has been added in Section 5.6.

Reviewer 4:

The tasks in Task Bench are idealized compute or memory-bound loops; problem size is simulated by controlling the number of iterations in these loops.

Shared memory is permitted and is used in our MPI+OpenMP implementation (among others).

The code for determining the dependence pattern and implementations of
kernels are contained in the core. A code sample from an
implementation is included in Listing 2. As the reviewer can see,
implementations can be quite simple.

Reviewer 5:

As described in Section 6, citations [27] and [37] are the only papers
we're aware of which provide an empirical performance evaluation of a
substantial number of systems. Many mini-app and benchmark papers only
include a small number of systems, or describe systems but do not
include an empirical evalution of performance. When there are
follow-up papers which describe additional implementations, they
rarely make comparisons beyond the reference implementation of the
code.

The output of a task is an array of bytes overwritten with a repeating
pattern (timestep, point). If the input does not match an assertion is
thrown.

### 2019-05-23 0_abstract.tex
Update word count.
Update list of systems.
Only claim 15 systems so we don't have to quibble about MPI+CUDA not running on Cori.
Apply Alex's comments.

\begin{abstract}
% 128 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench seeks to provide the ease of implementation and portability of
microbenchmarks, which makes comparisons between many systems
possible, with the ability to model a wide range of application
scenarios, giving it coverage approximating mini-apps.

We conduct a comprehensive study with implementations of Task Bench in
13 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce and use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of different dependence patterns. We also
study each system's scalability, ability to hide communication and
mitigate load imbalance.
\end{abstract}

>>>>

\begin{abstract}
% 147 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: all benchmarks constructed with Task Bench automatically
run on all Task Bench implementations, and vice versa. Furthermore,
Task Bench's parameterization allows a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

We conduct a comprehensive study with implementations of Task Bench in
15 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce and use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of different dependence patterns. We also
study each system's scalability, ability to hide communication and
mitigate load imbalance.
\end{abstract}

### 2019-05-23 1_introduction.tex
Apply Alex's comments.

One approach to comparing the performance of different systems is through \emph{proxy-} or
\emph{mini-apps}. Because they distill the key computational characteristics of a larger
application, mini-apps can provide insight
without the expense of developing a production code. However, despite the name, our experience is that
mini-apps still require significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop a correct implementation. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Few published studies involve more than a handful of systems~\cite{LULESH13}.

@@

To address the first issue, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative studies.

>>>>

One approach to comparing the performance of different systems is through \emph{proxy-} or
\emph{mini-apps}. Because they distill the key computational characteristics of a larger
application, mini-apps can provide insight
without the expense of developing a production code. However, despite the name, our experience is that
mini-apps still require significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop a correct implementation. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Few published studies compare more than a handful of systems~\cite{LULESH13}.

@@

To address the first issue, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative
studies. Additionally, Task Bench is separated into a core API shared
by all the system implementations. New benchmarks immediately run on
all systems, and new systems immediately run all benchmarks included
in Task Bench. Thus, Task Bench is much more productive for
benchmarking than the standard approach in which each benchmark must
be implemented separately for each system.

### 2019-05-23 3_task_bench.tex
Add kernel code sample.
Tweak text.
Apply Alex's comments.

To measure a wide variety of
systems and application scenarios, we developed a novel
parameterized benchmark called Task Bench. Task Bench
implementations execute a \emph{task graph}, with tasks for each point
in an \emph{iteration space} and dependencies between tasks determined
by a \emph{dependence relation}. A task that depends on another task
must execute after that task, and receives the output data from that
task as input. Each task can execute any one of a
number of kernels (compute-bound, memory-bound, etc.) and can generate
a configurable amount of output to control the amount of communication
performed with each task dependency. Thus Task Bench can explore a large space of application scenarios to
understand the performance of parallel and
distributed programming systems.

@@

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. The central aspects of Task Bench, such as generating
task graphs and enumerating dependencies, are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all systems, eliminating a potential
source of performance disparity that can be a pitfall for
implementations of mini-apps. Finally, the Task Bench core library manages the
parsing of input parameters and the display of output results,
ensuring that all implementations behave in a uniform manner and can
be scripted consistently. Because much of the functionality needed for
a Task Bench implementation is in the core library, implementations of
Task Bench are small: Our 15 Task Bench implementations range from 88
to 1500 lines, with several hundred lines being typical.

\input{f5_task_graphs}
\input{t0_equations}
\input{l1_code_sample}

@@

>>>>

To measure a wide variety of
systems and application scenarios, we developed a novel
parameterized benchmark called Task Bench. Task Bench
implementations execute a \emph{task graph}, with tasks for each point
in an \emph{iteration space} and dependencies between tasks determined
by a \emph{dependence relation}. A task $A$ that depends on another task $B$
must execute after task $B$, and receives the output data from
task $B$ as input. Each task can execute any one of a
number of kernels (compute-bound, memory-bound, etc.) and can generate
a configurable amount of output to control the amount of communication
performed with each task dependency. Thus Task Bench can explore a large space of application scenarios to
understand the performance of parallel and
distributed programming systems.

@@

\input{f5_task_graphs}
\input{t0_equations}

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. The central aspects of Task Bench, such as generating
task graphs and enumerating dependencies, are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all systems, eliminating a potential
source of performance disparity that can be a pitfall for
implementations of mini-apps. Finally, the Task Bench core library manages the
parsing of input parameters and the display of output results,
ensuring that all implementations behave in a uniform manner and can
be scripted consistently. Because much of the functionality needed for
a Task Bench implementation is in the core library, implementations of
Task Bench are small: Our 15 Task Bench implementations range from 88
to 1500 lines, with several hundred lines being
typical. Listings~\ref{lst:compute-kernel} and \ref{lst:code-sample}
show excerpts from the Task Bench core, and the Dask implementation,
respectively. Only the code in \ref{lst:code-sample} need be
implemented for each system, minimizing the work required for each additional system.

@@

\input{l0_compute_kernel}
\input{l1_code_sample}

### 2019-05-23 5_evaluation.tex
Apply Alex's comments.

For GPU experiments we use the Piz Daint supercomputer~\cite{PizDaint}, a Cray XC50 with
one Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
P100 per node. We use GCC \textbf{X.Y}, Cray MPICH \textbf{Z.W}, and
CUDA \textbf{A.B}.

@@

\subsection{Compute-Bound Performance}
\label{subsec:peak-performance-and-efficiency}

@@

\subsection{Memory-Bound Performance}

@@

Not all cores are required to saturate memory bandwidth. As a result,
systems that reserve one or more cores are able to hit 100\% of peak
performance, unlike with compute-bound kernels. Otherwise the
performance curves are similar. We are currently investigating the
performance of Charm++ and Regent.

% FIXME: why are Charm++ and Regent slow?

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency.

@@

Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which compares favorably with the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the x-axis is measured in the number of FLOP
performed during the run, as the GPU kernels must perform more work to
achieve peak performance.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus achieve
some overlap between communication and computation. We see that the
overdecomposed \lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities. Thus at smaller problem sizes the CPU achieves higher
performance.

>>>>

For GPU experiments we use the Piz Daint supercomputer~\cite{PizDaint}, a Cray XC50 with
one Intel Xeon E5-2690 v3 (12 physical cores) and one NVIDIA Tesla
P100 per node. We use GCC 6.2.0, Cray MPICH 7.7.2, and
CUDA 9.1.85.

@@

\subsection{Compute Kernel Performance}
\label{subsec:peak-performance-and-efficiency}

@@

\subsection{Memory Kernel Performance}

@@

Not all cores are required to saturate memory bandwidth. As a result,
systems that reserve one or more cores are able to hit 100\% of peak
performance, unlike with compute-bound kernels. Otherwise the
performance curves are similar. We are currently investigating the
performance of Charm++, Regent and X10.

% FIXME: why are Charm++, Regent and X10 slow?

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency. An interesting observation is that
the systems with the smallest METG on one node have roughly an order
of magnitude higher METG at 256 nodes. The obvious conclusion is that
increased communication latency requires significantly larger tasks to
achieve the same level of efficiency.

@@

Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which compares favorably with the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the x-axis is measured in the number of FLOPs
performed during the run, as the GPU kernels must be given more work to
achieve peak performance.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus achieve
some overlap between communication and computation. We see that the
overdecomposed \lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities. Thus at smaller problem sizes the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of MTEG (as the peak performance of the GPU and CPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: The cost of communicating data and
tasks to and from the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage of
MPI's very lightweight mechanisms.

### 2019-05-23 6_related.tex
Apply Alex's comments.

Programming systems for parallel and distributed computing are most
commonly evaluated using proxy- or mini-apps, or
micro\-benchmarks. Mini-apps are explicitly derived from larger
applications and therefore have the advantage of bearing some
relationship to the original. This advantage typically does not hold
for microbenchmarks.

Though smaller than full applications, mini-apps can be challenging to
implement to a level of quality sufficient for conducting comparative
studies between programming systems. The largest study we know of,
for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI). Other
studies usually lack a comprehensive evaluation, even if multiple
implementations are available:

@@

Microbenchmarks are often easier to implement, but may not be
representative of real-world applications. From the PRK
suite~\cite{PRK14}, the PRK Stencil code has been reasonably widely
implemented, and a follow-up paper~\cite{PRKRuntimes16} compares MPI
variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
single, simple code involving 2 kernels (which combined fit in under 50 lines
of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} are also relatively widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. The benchmarks consist mostly of small
kernels for dense array computations. Neither the PRK nor the NAS
benchmark suite achieve the breadth of coverage, flexibility, or ease
of implementation of Task Bench. We believe the evidence for ease of
implementation is clear; Task Bench is implemented for more systems
than all previous mini-apps and microbenchmarks. Task Bench is also
not a fixed set of benchmarks, and it is even easier to use Task Bench
to generate a new benchmark to measure some aspect of system
performance.

>>>>

Parallel and distributed programming systems are most
commonly evaluated using proxy- or mini-apps, or
micro\-benchmarks. Mini-apps are explicitly derived from larger
applications and therefore have the advantage of bearing some
relationship to the original. This advantage typically does not hold
for microbenchmarks.

Though smaller than full applications, mini-apps can be challenging to
implement to a level of quality sufficient for conducting comparative
studies between programming systems. The largest study we know of,
for the mini-app LULESH~\cite{LULESH13}, compares 7 programming
systems (Chapel, Charm++, CUDA, Lizst, Loci, OpenMP, and MPI), each of
which require a wholly separate, tuned implementation (in contrast to
Task Bench). Other
studies usually lack a comprehensive evaluation, even if multiple
implementations are available:

@@

Microbenchmarks are often easier to implement, but may not be
representative of real-world applications. From the PRK
suite~\cite{PRK14}, the PRK Stencil code has been reasonably widely
implemented, and a follow-up paper~\cite{PRKRuntimes16} compares MPI
variants, SHMEM, UPC, Charm++, and Grappa. However, PRK Stencil is a
single, simple code involving 2 kernels (which combined fit in under 50 lines
of C++) and a halo exchange on a structured grid. The NAS benchmark suite~\cite{NAS91, NAS95} is also relatively widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. The benchmarks consist mostly of small
kernels for dense array computations. Neither the PRK nor the NAS
benchmark suite achieve the breadth of coverage, flexibility, or ease
of implementation of Task Bench. We believe the evidence for ease of
implementation is clear; Task Bench is implemented for more systems
than all previous mini-apps and microbenchmarks. Task Bench is also
not a fixed set of benchmarks, and it is even easier to use Task Bench
to generate a new benchmark to measure some aspect of system
performance.

### 2019-05-23 7_conclusion.tex
Edits from Alex.
Apply Alex's comments.

Task Bench is a promising new approach for evaluating the performance
of parallel and distributed programming systems. We have seen that
systems vary by orders of magnitude in runtime overhead, as measured
by METG. Widely used systems in industry have METG measured in
seconds, reflecting the very coarse tasks and lack of need for strong
scaling performance in current workloads. At the other extreme,
traditional HPC systems have METG in the range of 10-100 \textmu{}s,
necessary for strong scaling to small task granularities. The newer
task-based systems, with METG in the range of 100 \textmu{}s to 1 ms
have performance sufficient for weak scaling many HPC workloads, but
more work is needed to strong scale the most demanding applications.

>>>>

Task Bench is a new approach for evaluating the performance
of parallel and distributed programming systems. We have seen that
systems vary by orders of magnitude in runtime overhead, as measured
by METG. Widely used systems in industry have METG measured in
seconds, reflecting the very coarse tasks and lack of need for strong
scaling performance in current workloads. At the other extreme,
traditional HPC systems have METG in the range of 10-100 \textmu{}s,
necessary for strong scaling to small task granularities, though we
have shown that METG also increases significantly for these systems
with the size of the cluster and the use offload accelerators. The newer
task-based systems, with METG in the range of 100 \textmu{}s to 1 ms
have performance sufficient for weak scaling many HPC workloads, but
more work is needed to strong scale the most demanding applications.
